---
output:
  md_document:
    variant: markdown_github+backtick_code_blocks
---

```{r options, echo = F}
opts_chunk$set(fig.path = "docs/figures/", comment = NA, tidy = F, message = F, warning = F, echo = F, results = 'verbatim')
```

```{r inc1}
library(ggplot2)
library(pander)
```

```{r inc2}
library(devtools)
load_all()
```

```{r settings}
theme_set(theme_light())
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
```

# athaliana

## About

`athaliana` is an R data package for the A. thaliana data set https://github.com/Gregor-Mendel-Institute/atpolydb. 

## Code examples

### Get necessary data with a few functions

Here we are inerested to run a polygenic model for one of the traits (`FRI`).

* The necessary data are the table of phenotypes & the relatedness matrix (pre-calculated from the SNPs data, see `athaliana_compute_relmat` function).
* The mixed model comes from the `lme4qtl` R package.

```{r poly, echo = T}
phen <- athaliana_phen(traits = "FRI")
relmat <- athaliana_relmat()

library(lme4qtl)
m <- relmatLmer(FRI ~ (1|id), phen, relmat = list(id = relmat))

m
```
