---
title: "GWAS on the A. thialana data set"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{GWAS on the A. thialana data set}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Settings

### Parameters

```{r par}
chr <- 4

cores <- 2
```

### Packages

```{r inc}
library(athaliana)

library(dplyr)

library(lme4qtl)
```

## Load the data

```{r data, cache = T}
phen <- athaliana_phen(traits = "FRI", rows_order = "snp")
relmat <- athaliana_relmat()
```

```{r gdata, cache = TRUE}
gdat <- athaliana_snp(chr = 4)
  
gdat <- bind_cols(phen["FRI"], gdat)

gdat_subset <- gdat[1:20]
```

## Polygenic model

```{r poly, cache = T}
(poly <- relmatLmer(FRI ~ (1|id), phen, relmat = list(id = relmat)))
```

## Association study of a few snps

```{r assoc, cache = T}
snps <- names(gdat_subset) %>% grep("^snp", ., value = TRUE)

assoc <- assocLmer(FRI ~ (1|id), gdat_subset, relmat = list(id = relmat), 
  data_snp_cov = subset(gdat_subset, select = snps), 
  batch_size = 10, cores = cores)
```


## GWAS

```{r gwas, eval = F}
snps <- names(gdat) %>% grep("^snp", ., value = TRUE)

gwas <- assocLmer(FRI ~ (1|id), phen, relmat = list(id = relmat), 
  data_snp_cov = gdat[-1],
  batch_size = 100, cores = cores)
```


## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
